A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-08-04 demystify.tex
updates

In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates. Also, most CS programs already receive many applications, and would be overwhelmed if the application process were free.

@@

\paragraph{Why Self-written Letters are Bad?} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs. Such letters have \textbf{very little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?  Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter). Experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.

\paragraph{Why Generic Letters are Bad?} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak. This does not mean the professor is not good or does not care about you, but they just do not know you well enough to write a strong letter.
So it might be a good idea to directly ask if the prof. is willing to write a \emph{strong} letter for you. If not, then you should ask someone else.  For example, if a student I don't know well asks me to write a letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their case.

@@

\paragraph{Why You Should Waive Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications. When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words. It's also about keeping things open and honest between you and your letter writers and encourages them to be real about your strengths and qualifications. Plus, it keeps things private.

@@

\paragraph{Why You Want to Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value. One thing you can do to help your writer to write a strong, customized LoR for you is providing details or unique things about yourself. For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention. You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
Sometimes your writer will explicitly ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).

\paragraph{Why You Need to Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.  You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.  Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).  Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid-January).

@@

In your statement, you have the opportunity to connect directly with us and make your application stand out and unique. Your SOP can convince adcom members that you would \emph{fit} the program, even if you don't have very strong research experience.
A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
% is important because if you need (GTA) funding, it will provide evidence
% that you can teach and communicate with students.

@@

\textbf{Advice}: If your school is the top in your country but you think it might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that). Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}. %It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.

@@

\textbf{Advice}: As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
Note that while having good grades might not help much,
having bad grades (e.g., < 3.0) will be \red{red flag} (unless your LORs or
statement give a proper explanation). This is especially true if you
have bad grades in relevant courses, e.g., Math and CS.

@@

\paragraph{GRE} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it. Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it. It does not help much!

@@

Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.  For example, at Mason also waive the English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.

\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., plagiarism, cheating, or other academic dishonesty). Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time). Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.

@@

\paragraph{Rejection Letters} Typically, schools start sending out rejection letters to remaining applicants after they have finalized their admissions decisions. Thus, rejection letters are often sent out late (e.g., after April 15th or even much later). Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks). In short, you just have to be patient and wait.

>>>>

In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates. Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.

@@

\paragraph{Self-written letters are bad} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs. Such letters have \textbf{very little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?  Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter). Experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.

\paragraph{Generic letters are bad} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak. This does not mean the professor is not good or does not care about you, but they just do not know you well enough to write a strong letter.
So it might be a good idea to directly ask if the prof. is willing to write a \emph{strong} letter for you. If not, then you should ask someone else.  For example, if a student I don't know well asks me to write a letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their case.

@@

\paragraph{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications. When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words. It's also about keeping things open and honest between you and your letter writers and encourages them to be real about your strengths and qualifications. Plus, it keeps things private.

@@

\paragraph{Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value. One thing you can do to help your writer to write a strong, customized LoR for you is to provide details or unique things about yourself. For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention. You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
Sometimes your writer will explicitly ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).

\paragraph{Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.  You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.  Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).  Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid-January).

@@

In your statement, you have the opportunity to connect directly with adcom members and make your application stand out and unique. Your SOP can convince adcom members that you would \emph{fit} the program, even if you don't have very strong research experience.
A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
% is important because if you need (GTA) funding, it will provide evidence
% that you can teach and communicate with students.

@@

If your school is the top in your country but you think it might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that). Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}. %It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.

@@

As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
Note that while having good grades might not help much,
having bad grades (e.g., < 3.0) will be \red{red flag} (unless your LORs or
statement give a proper explanation). This is especially true if you
have bad grades in relevant courses, e.g., Math and CS.

@@

\paragraph{GREs Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it. Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it. It does not help much!

@@

Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.  For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.

\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdraw grades, plagiarism, cheating, or other academic dishonesty). Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time). Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.

@@

\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}. Thus, rejection letters are often sent out late (e.g., after April 15th or even much later). Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks). In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).

### 2024-08-04 faculty.tex
updates

    % Title
    \node[align=center] at ($(current page.center)+(0,-5)$) 
    {
    {\fontsize{24}{24} \selectfont {{Demystifying Tenure Track:}}} \\[0.15in]
    {\fontsize{18}{18} \selectfont {{Navigating the Faculty Journey In Computer Science}}} \\[1in]    
    %{\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]

@@

\chapter{Research}
Research here focuses on paper publication, which is a major component in getting your name out there and of course in your departmental evaluation.

@@

@@

\subsection{Internal Funding}
These are funds from your university, city, or state (e.g., Virginia has the CCI program that provide ``seed'' grants for researchers from universities in Virginia).  Your ``startup'' package is also be considered as a large funding from an internal source (your university) to kickstart your research. These funding are less competitive than NSF, but they are also less prestigious (e.g., they aren't much and usually don't count as much for tenure).  However, they are good sources of funding to get preliminary results and later apply for traditional grants (and many internal funding programs explicitly encourage this).

\subsection{Industrial Gifts}
Industrial gifts are \emph{unrestricted funding} from companies such as Amazon, Microsoft, Facebook, and Google. These are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).  They are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., summary salary, equipments, or even bowling events for your group).

\paragraph{Personal Experiences}
As with most tenure track professors, I focus mostly on NSF grants and most of my funding comes from NSF.  I only have one grant from Army Research Office (ARO) in which I am a co-PI and didn't have to write the proposal. I have tried to get other DoD funding, but without much success, I found that DoD program managers are very picky and often have their own agenda.  I have also received internal funding from my university and industrial gifts from Amazon,and Facebook.  I have found that industrial gifts are the best kind of funding, as they are unrestricted and have no associated timeline by when you should spend them.  They are also highly competitive and prestigious, and are a great way to get funding for your research.

@@

\chapter{Teaching}
\section{Teaching Load}
\section{Teaching Evaluation}

>>>>

    % Title
    \node[align=center] at ($(current page.center)+(0,-5)$) 
    {
    {\fontsize{24}{24} \selectfont {{Demystifying Tenure Track:}}} \\[0.15in]
    {\fontsize{18}{18} \selectfont {{My Academic Journey in Computer Science}}} \\[1in]    
    %{\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]

@@

\chapter{Research}
This section focuses on paper publication, which is a major component in getting your name out there and of course in your departmental evaluation.

\section{What is needed for tenure?}

Publications are a major component of tenure evaluation. There are no specific numbers of papers or which conferences/journals are required, but typically you need to have a good number of papers in top venues to get tenure.  Also work that won awards (e.g., best paper awards or 10-year most influential awards) also count a lot in tenure evaluation. Citations are not required but work with very high citations would help.

% 5.1 Evaluating Research from the Point of View of Tenure

% 	•	Key Components: Publication record and expert letters.
% 	•	Consistency & Impact: Steady publication in top venues, with standout work defining one’s reputation.
% 	•	Research Independence: Importance of showing independent work, separate from past advisors.

% 5.2 Coming up with a Research Agenda

% 	•	Research Agenda vs. Project: A broad goal guiding multiple projects.
% 	•	Criteria: Inspiring, technically challenging, and doable work, ideally tied to the professor’s expertise.

% 5.3 Picking Research Problems

% 	•	Importance of Selection: Prioritizing meaningful, high-impact problems.
% 	•	Criteria: Should be either intellectually elegant or practically useful, balancing theory and practical applications.

% 5.4 Knowing When to Give Up

% 	•	Decision Making: Assessing progress and potential, setting early criteria for success, and being willing to pivot if necessary.

% 5.5 Managing Collaborations

% 	•	Balancing Collaborations: Limiting them to maintain focus on primary research.
% 	•	Industrial Collaborations: Handling agreements, IP, and publication protocols.

% The chapter covers developing a research agenda, selecting impactful problems, knowing when to pivot, and managing collaborations, emphasizing focus and strategic decision-making.

@@

\section{From a Tenure Evaluation Point of View}
Without a doubt, grants are one of the most, if not the most, important component of tenure evaluation.  The most common reason for not getting tenure is not having enough funding. In an ideal world, research results, e.g., publications, would be the more important; but in reality, funding is often considered more important for tenure evaluation at most R1 universities.
The amount of funding you have is a direct reflection of your research impact and productivity.  It is also a reflection of your ability to write proposals, manage projects, and work with collaborators. More realistically, grants are needed to pay for your students, travel conference, research equipments, and even your own summer salary.

In general, the more funding you have, the better.  However, the quality of the funding is also important.  For example, external NSF grants are more prestigious than an internal grant.  Also, funding from industry, while has a much smaller amount, is often highly competitive, prestigious, and can help show that your research is of interest to industry.  In general, you should aim to get funding from a variety of sources, including NSF, Defense, industry, and internal grants.  This will show that you are able to work with different types of sponsors and that your research is of interest to a wide range of people.

@@

\subsection{Internal Funding}
These are funds from your university, city, or state (e.g., Virginia has the CCI program that provide ``seed'' grants for researchers from universities in Virginia).  Your ``startup'' package is also be considered as a large funding from an internal source (your university) to kickstart your research. These funding are less competitive than NSF, but they are also less prestigious (e.g., they aren't much and usually don't count as much for tenure).  However, they are good sources of funding to get preliminary results and later apply for traditional grants.

\subsection{Industrial Gifts}
Industrial gifts are \emph{unrestricted funding} from companies such as Amazon, Microsoft, Facebook, and Google. While some are connection-based (e.g., for your ongoing work with a collaborator at Facebook), most are based on the merits of your research.
These are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).  
They also show that your research has values and impact beyond academia and is interested to industry. Practically, they are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., summary salary, equipments, or even bowling events for your group).

\paragraph{Personal Experiences}
As with most tenure track professors, I focus mostly on NSF grants and most of my funding comes from NSF.  I only have one grant from Army Research Office (ARO) in which I am a co-PI and didn't have to write the proposal. I have tried to get other DoD funding, but without much success, I found that DoD program managers are very picky and often have their own agenda.  I have also received internal funding from my university and industrial gifts from Amazon, and Facebook.  I have found that industrial gifts are the best kind of funding, as they are unrestricted and have no associated timeline by when you should spend them.  They are also highly competitive and prestigious, and are a great way to show research impact.

@@

\chapter{Teaching}

\section{From a P\&T Point of View}

While teaching should take about 40\% of your time and your salary is mainly for teaching, teaching is not as important as research for tenure.  Moreover, for most new faculty, teaching is not as difficult as getting grants. Thus, tenure-track faculty often do not worry much about teaching.

\section{Teaching Load}
\section{Teaching Evaluation}
